---
title: "Assignment 2 - Logistic Regression"
author: "ITAO 40420 - Machine Learning"
date: "29 January 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assignment 2

For this analysis we will be using churn data. The dataset is stored as `churn_data.rda` which can be loaded in by running the below command if the file is located in your working directory or else by accessing the file through file->open file

```{r Load Data 2}
load("churn_data.rda") # Load churn dataset
```

We can view a summary of this dataset by running `summary()`:

```{r}
summary(churn_data)
```

The variables we have are:

* Gender - Whether the customer is a male or a female
* SeniorCitizen - Whether the customer is a senior citizen or not (1,0)
* Partner - Whether the customer has a partner or nor (Yes, No)
* Dependents - Whether the customer has dependents or not (Yes, No)
* tenure - Number of months the customer has stayed with the company
* PhoneService - Whether the customer has a phone service of no (Yes, No)
* MultipleLines - Whether the customer has multiple lines or not (Yes, No, No phone service)
* InternetService - Customer's internet service provider (DSL, Fiber optic, No)
* OnlineSecurity - Whether the customer has online security or not (Yes, No, No internet service)
* OnlineBackup - Whether the customer has online backup or not (Yes, No, No internet service)
* DeviceProtection - Whether the customer has tech support or not (Yes, No, No internet service)
* StreamingTV - Whether the customer has streaming TV or not (Yes, No, No internet service)
* StreamingMovies - Whether the customer has streaming movies or not (Yes, No, No internet service)
* Contract - The contract term of the customer (Month-to-month, One year, Two year)
* PaperlessBilling - Whether the customer has paperless billing or not (Yes, No)
* PaymentMethod - The customer's payment method (Electronic CHECK, Mailed check, Bank transfer (automatic), Credit card (automatic))
* MonthlyCharges - The amount charged to the customer monthly
* TotalCharges - The total amount charged to the customer
* Churn - Whether the customer churned or not (1 = Yes or 0 = No)


The response variable for this dataset is `Churn`:

```{r}
summary(as.factor(churn_data$Churn))
```


For the assignment please carry out the following tasks/answer the following questions: (12 Points)


* 1 - Create at least two visualizations potentially revealing factors with predictive power (2 Points)

```{r}
library(ggplot2)
plot_dat <- churn_data # Create data to use as plot data
plot_dat$Churn <- as.factor(plot_dat$Churn) # Convert response to factor for plotting
g_1 <- ggplot(plot_dat, aes(x = gender, fill = Churn)) + # Set x as mean radius and fill as diagnosis
  geom_bar(alpha = 0.5, position = "dodge") + # Select density plot and set transperancy (alpha)
    theme_set(theme_bw(base_size = 22) ) + # Set theme and text size
  theme(panel.grid.major = element_blank(), # Remove grid
        panel.grid.minor = element_blank(), # Remove grid
        panel.border = element_blank(), # Remove grid
        panel.background = element_blank()) + # Remove grid 
  labs(x = "Gender", title = "Gender - Churn or Not",      
       fill = "Churn") + # Set labels
  scale_fill_manual(values = c("1" = "red", "0" = "blue"), # Set fill colors manually
                    labels = c("1" = "Churn", "0" = "Not Churn")) # Set labels for fill
g_1 # Generate plot
```
```{r}
table(churn_data$PaymentMethod)

library(ggplot2)
plot_dat <- churn_data # Create data to use as plot data
plot_dat$Churn <- as.factor(plot_dat$Churn) # Convert response to factor for plotting
g_2 <- ggplot(plot_dat, aes(x = PaymentMethod, fill = Churn)) + # Set x as mean radius and fill as diagnosis
  geom_bar(alpha = 0.5, position = "dodge") + # Select density plot and set transperancy (alpha)
    theme_set(theme_bw(base_size = 8) ) + # Set theme and text size
  theme(panel.grid.major = element_blank(), # Remove grid
        panel.grid.minor = element_blank(), # Remove grid
        panel.border = element_blank(), # Remove grid
        panel.background = element_blank()) + # Remove grid 
  labs(x = "PaymentMethod", title = "PaymentMethod - Churn or Not",      
       fill = "Churn") + # Set labels
  scale_fill_manual(values = c("1" = "red", "0" = "blue"), # Set fill colors manually
                    labels = c("1" = "Churn", "0" = "Not Churn")) # Set labels for fill
g_2 # Generate plot
```



* 2 - Fit a logistic regression model to the data. (Remember you can use `resp ~.,` to include all explanatory variables in the model) (2 Points)

```{r}
library(glmnet) # Load glmnet
fit_1 <- glm(Churn ~., # Set formula
             family=binomial(link='logit'), # Set logistic regression
             data= churn_data) # Set dataset
sum_1 <- summary(fit_1) # Sumamrize model
sum_1
```


* 3 - What features are significant in your logistic regression model? (1 Point)

```{r}
sum_1$coefficients[sum_1$coefficients[,4] < 0.01, c(1, 4)]
```


* 4 - What is the AIC of your logistic regression model?  (1 Point)
The AIC of my model is 5874.3.

* 5 - Add an interaction term to the logistic regression model? (1 Point)

```{r}
fit_2 <- glm(Churn ~. + Contract*PaperlessBilling, # Set formula
             family=binomial(link='logit'), # Set logistic regression
             data= churn_data) # Set dataset
sum_2 <- summary(fit_2) # Sumamrize model
sum_2
```


* 6 - Does the addition of the interaction term increase or decrease the AIC? What is the new AIC for the model? (1 Point)
Yes, it increases a little bit. The new AIC is 5877.8


* 7 - Fit a lasso logistic regression model to the data. (1 Point)

```{r}
churn_data[, c(5, 18, 19)] <- scale(churn_data[, c(5, 18, 19)])

x_vars <- model.matrix(Churn ~., 
                       churn_data)[,-1]


fit_3 <- glmnet(x = x_vars, # Fit explanatory variables
                y = churn_data$Churn, # Fit response variable
                alpha = 1, # Set alpha as 1 for lasso
                family = "binomial",
                lambda = 0.05) 

coef(fit_3)
```


* 8 - Which coefficients were non-zero in the lasso model? Were they positive or negative? What does this imply? (1 Point)

Coefficients of tenure, InternetServiceFiber optic, PaymentMethodElectronic check are all non-zero in the lasso model. 
Coefficients of InternetServiceFiber optic, PaymentMethodElectronic check are positive, and coefficient of tenure is negative.
This implies that InternetServiceFiber optic, PaymentMethodElectronic would make the outcome: probability of churn rises significantly, and tenure would make the outcome: probability of churn decreases significantly.


2 Points for analysis decisions and code quality.

Please submit an an R-markdown document with the answers, optionally you can submit an R-script with the code used for your analysis and a document answering the specific questions, you can use word, pdf etc.

Assignments can be submitted via Canvas or by email to mbarron2@nd.edu.







##For cross validation (fold=k) in cp, is it mean for new sample fit cv or lambda decision cp? (decision tree) (rel_error means new_sample vs train_sample?)

##Why full tree cp = 0? (how about lambda*splits?)

